---
title: "deleted variable exploration"
author: "Yiru Gong, yg2832"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r}
library(tidyverse)
library(Rtsne)
library(cluster)
```

import data:
* new_full: all data with na < 30000, 868 cols
* final: final enlarged dataset, 260 cols

```{r}
load("final_na_enlarged.Rdata")
unused_idx = setdiff(colnames(new_full),colnames(final))
unused = new_full[,unused_idx]
# summary(unused[,1:10])
write.csv(unused, "unused.csv",row.names = F)
```

```{r}
tsne = Rtsne(unused)
tsne_df = tsne$Y %>% 
  as.data.frame() %>% 
  rename(tsne1 = "V1",
         tsne2 = "V2")
tsne_df = cbind(tsne_df, hi_flag = final$hi_flag, cat)
```

```{r}
pam = pam(unused,10)


```


```{r}
heading = read_excel('../Humana_Mays_2022_DataDictionary.xlsx', sheet = 'Data Dictionary')
# heading = heading %>% janitor::clean_names()
heading["used"] = sapply(heading$`Feature name`,function(x) ifelse(x %in% colnames(final), "yes","no"))

setdiff(colnames(final),heading$`Feature name`[which(heading$used=='yes')])
write.csv(heading,"labeled_dictionary.csv",row.names = F)
```


## added manually by Xingran

```{r}
added = c("atlas_naturalchangerate1016",'atlas_net_international_migration_rate',
'atlas_snapspth16',
'cmsd1_ano_pmpm_ct',
'cmsd1_ext_pmpm_ct',
'cmsd1_neo_pmpm_ct',
'cmsd1_pre_pmpm_ct',
'cmsd1_unc_pmpm_ct',
'cons_ccip')

added[added %in% colnames(final)]
added[! added %in% colnames(final)]

load("final.Rdata")
final_small = final
load("final_na_enlarged.Rdata")
final_large = final
final_manual = cbind(final_small,
                     final_large[,added[added %in% colnames(final)]],
                     new_full[,added[! added %in% colnames(final)]])
save(final_manual, rowTrain, cat_idx, file = "to_xinran.Rdata")
```

